### 2024-08-30 download-football-data.qmd
20240830 - merging

Below, we provide examples for how to download various types of NFL data.
For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
We save each data file after downloading it, so we can use the data in subsequent chapters.
Guidance for how to merge the various data files is provided at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md>

>>>>

Below, we provide examples for how to download various types of NFL data.
For additional resources, @Congelio2023 provides a helpful introductory text for working with NFL data in `R`.
We save each data file after downloading it, so we can use the data in subsequent chapters.
This chapter uses extensive use of merging to process the data for later use.
See @sec-merging for a reminder of how to perform merging, the types of merges, and what you can expect when you merge data objects with different formats.
Guidance for how to merge the various NFL-related data files is provided at the following link: <https://github.com/nflverse/nfldata/blob/master/DATASETS.md>.

### 2024-08-30 getting-started.qmd
20240823 - git
20240830 - merging

1. Set up a free account on [GitHub.com](https://github.com).
1. Download GitHub Desktop: <https://desktop.github.com>
1. Make sure you are logged into your GitHub account on [GitHub.com](https://github.com).
1. Go to the following GitHub repository: <https://github.com/isaactpetersen/QuartoBlogFantasyFootball> and complete the following steps:
    1. Click "Use this Template" (in the top right of the screen) > "Create a new repository"
    1. Make sure the checkbox is selected for the following option: "Include all branches"
    1. Make sure your Owner account is selected
    1. Specify the repository name to whatever you want, such as `FantasyFootballBlog`
    1. Type a brief description, such as `Files for my fantasy football blog`
    1. Keep the repository public (this is necessary for generating your blog)
    1. Select "Create repository"
1. After creating the new repository, make sure you are on the page of of your new repository and complete the following steps:
    1. Click "Settings" (in the top of the screen)
    1. Click "Actions" (in the left sidebar) > "General"
    1. Make sure the following are selected:
        - "Read and write permissions" (under "Workflow permissions")
        - "Allow GitHub Actions to create and approve pull requests"
        - then click "Save"
    1. Click "Pages" (in the left sidebar)
    1. Make sure the following are selected:
        - "Deploy from a branch" (under "Source")
        - "gh-pages/(root)" (under "Branch")
        - then click "Save"
1. Clone the repository to your local computer by clicking "Code" > "Open with GitHub Desktop", select the folder where you want the repository to be saved on your local computer, and click "Clone"
:::

@@

Merging (also called joining) merges two data objects using a shared set of variables called "keys."
The keys are the variable(s) that uniquely identify each row (i.e., they account for the levels of nesting).
In some data objects, the key might be the player's identification number (e.g., `player_id`).
However, some data objects have multiple keys.
For instance, in long form data objects, each participant may have multiple rows corresponding to multiple seasons.
In this case, the keys may be `player_id` and `season`.
If a participant has multiple rows corresponding to seasons and games/weeks, the keys are `player_id`, `season`, and `week`.
In general, each row should have a value on each of the keys; there should be no missingness in the keys.

@@

@@

Here are the original data in long form.
The data are structured in "player-season-week form".
That is, every row in the dataset is uniquely identified by the combination of variables, `ID`, `season`, and `week`.
This is an example of long form, because each player has multiple rows.

@@

Below, we widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "player form" (where each row is uniquely identified by the `ID` variable):

>>>>

1. Download and install `git`: <https://git-scm.com/downloads>
1. Set up a free account on [GitHub.com](https://github.com).
1. Download and install `GitHub Desktop`: <https://desktop.github.com>
1. Make sure you are logged into your GitHub account on [GitHub.com](https://github.com).
1. Go to the following GitHub repository: <https://github.com/isaactpetersen/QuartoBlogFantasyFootball> and complete the following steps:
    1. Click "Use this Template" (in the top right of the screen) > "Create a new repository"
    1. Make sure the checkbox is selected for the following option: "Include all branches"
    1. Make sure your Owner account is selected
    1. Specify the repository name to whatever you want, such as `FantasyFootballBlog`
    1. Type a brief description, such as `Files for my fantasy football blog`
    1. Keep the repository public (this is necessary for generating your blog)
    1. Select "Create repository"
1. After creating the new repository, make sure you are on the page of of your new repository and complete the following steps:
    1. Click "Settings" (in the top of the screen)
    1. Click "Actions" (in the left sidebar) > "General"
    1. Make sure the following are selected:
        - "Read and write permissions" (under "Workflow permissions")
        - "Allow GitHub Actions to create and approve pull requests"
        - then click "Save"
    1. Click "Pages" (in the left sidebar)
    1. Make sure the following are selected:
        - "Deploy from a branch" (under "Source")
        - "gh-pages/(root)" (under "Branch")
        - then click "Save"
1. Clone the repository to your local computer by clicking "Code" > "Open with GitHub Desktop", select the folder where you want the repository to be saved on your local computer, and click "Clone"
:::

@@

Merging (also called joining) merges two data objects using a shared set of variables called "keys."
The keys are the variable(s) that uniquely identify each row (i.e., they account for the levels of nesting).
In some data objects, the key might be the player's identification number (e.g., `player_id`).
However, some data objects have multiple keys.
For instance, in long form data objects, each participant may have multiple rows corresponding to multiple seasons.
In this case, the keys may be `player_id` and `season`—that is, the data are in `player`-`season` form.
If a participant has multiple rows corresponding to seasons and games/weeks, the keys are `player_id`, `season`, and `week`—in this case, the data are in `player`-`season`-`week` form.
In general, each row should have a value on each of the keys; there should be no missingness in the keys.

@@

When two objects are merged that have different formats, the resulting data object inherits the format of the data object that has more levels of nesting.
For instance, consider that you want to merge two objects, object A and object B.
Object A is in `player` form and object B is in `player`-`season`-`week` form.
When you merge them, the resulting data object will be in `player`-`season`-`week` form.

@@

Here are the original data in long form.
The data are structured in "`player`-`season`-`week` form".
That is, every row in the dataset is uniquely identified by the combination of variables, `ID`, `season`, and `week`—these are the keys.
This is an example of long form, because each player has multiple rows.

@@

Below, we widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable):
